One-shot learning: From domain knowledge to action models

### 2019-02-20 doc/ijcai-19-deductivelearning/deductivelearning.tex
added pointer to roni's paper to ijcai19 paper on deductive learning

(:observations (on blockA blockB) (on blockB blockC))
  \end{verbatim}
  \end{tiny}
	\caption{\small Example of a two-state observationn for the learning \strips\ action models.}
	\label{fig:observation}
\end{figure}

\subsection{Domain-specific knowledge}
Our approach is to introduce {\em domain-specific knowledge} in the form of {\em state-constraints} to restrict further the space of possible schemata. For instance, in the {\em blocksworld} one can argue that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists of an action schema $\xi$ because, in this specific domain, a block cannot be on top of itself. The notion of {\em state-constraint} is very general and has been used in different areas of AI and for different purposes.  If we restrict ourselves to planning, {\em state-constraints} are abstractions for the compact specification of a set of states. For example, {\em state-constraints} in planning allow to specify the set of states where a given action is applicable, the set of states where a given {\em axiom} or {\em derived predicate} holds or the set of states that are considered goal states.

@@

\subsection{Pruning inconsistent action models with {\em domain-specific knowledge}}
For every {\em state-constraint} $\phi\in\Phi$ we could extend $\mathsf{apply_{\xi,\omega}}$ actions with a conditional effect $\{\neg\phi\}\rhd\{mode_{inval}\}$ that checks the consistency of the {\em state-constraints} $\phi$ at every state traversed by a solution to the compiled problem. Cheking arbitrary $\phi$ formulae can however be too expensive for current classical planners. Instead, our approach is to check {\em state-constraints} in the form of {\em schematic mutex}. To implement this checking we add new conditional effects to {\em insert} and {\em apply} actions of the classical planning compilation for the learning of \strips\ action models~\cite{aineto2018learning}. These new conditional effects capture when the programmed model is inconsistent with a {\em schematic mutex} in $\Phi$.

Figure~\ref{fig:ceffects} summarizes the new conditional effects added to the classical planning compilation for the learning of \strips\ action models from {\em schematic mutex}. Next we describe each of them in detail:
\begin{enumerate}
\item[1-3.] For every {\em schematic mutex} $\tup{p,q}$ s.t. both $p$ and $q$ belong to $\in{\mathcal I}_{\Psi,\xi}$ a conditional effect is added to the $\mathsf{insertPre_{p,\xi}}$ actions to ban the insertion of two preconditions that are {\em schematic mutex}. Likewise, two conditional effects are added to the $\mathsf{insertEff_{p,\xi}}$ actions to ban the insertion of two positive/negative effects that are {\em schematic mutex}.
\item[4-5.] For every {\em schematic mutex} $\tup{p,q}$ s.t. both $p$ and $q$ belong to $\in{\mathcal I}_{\Psi,\xi}$ two conditional effects are added to the $\mathsf{apply_{\xi,\omega}}$ actions to ban positive effects that are inconsistent with an input observation.
\end{enumerate}

\begin{figure}
\begin{tabular}{lll}
{\bf ID} & {\bf Action} & {\bf New conditional effect}\\\hline
1.& $\mathsf{insertPre_{p,\xi}}$ & $\{pre\_q\_\xi\}\rhd\{mode_{inval}\}$\\
2.& $\mathsf{insertEff_{p,\xi}}$ & $\{pre\_q\_\xi,eff\_q\_\xi,pre\_p\_\xi\}\rhd\{mode_{inval}\}$\\
3.& $\mathsf{insertEff_{p,\xi}}$ & $\{\neg pre\_q\_\xi,eff\_q\_\xi,\neg pre\_p\_\xi\}\rhd\{mode_{inval}\}$\\
4.& $\mathsf{apply_{\xi,\omega}}$ & $\{\neg pre\_p\_\xi \wedge eff\_p\_\xi \wedge $\\
& & $q(\omega)\wedge \neg pre\_q\_\xi\}\rhd\{mode_{inval}\}$\\
5.& $\mathsf{apply_{\xi,\omega}}$ & $\{\neg pre\_p\_\xi \wedge eff\_p\_\xi \wedge $\\
& &$q(\omega)\wedge pre\_q\_\xi \wedge \neg eff\_q\_\xi\}\rhd\{mode_{inval}\}$
\end{tabular}
	\caption{\small Summary of the new conditional effects added to the classical planning compilation for the learning of \strips\ action models.}
	\label{fig:ceffects}
\end{figure}

The goals of the classical planning problem output by the original compilation are extended with the $\neg mode_{inval}$ literal to validate that only states {\em consistent} with the state constraints defined in $\Phi$ are traversed by the solution plans. This allows us to introduce a more compact definition $\mathsf{apply_{\xi,\omega}}$ actions than the one previously proposed by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning} that do not require disjunctions to code the possible preconditions of an action schema.

>>>>

(:observation (on blockA blockB) (on blockB blockC))
  \end{verbatim}
  \end{tiny}
	\caption{\small Example of a two-state observationn for the learning \strips\ action models.}
	\label{fig:observation}
\end{figure}

\subsection{The domain-specific knowledge}
Our approach is to introduce {\em domain-specific knowledge} in the form of {\em state-constraints} to restrict further the space of possible schemata. For instance, in the {\em blocksworld} one can argue that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists of an action schema $\xi$ because, in this specific domain, a block cannot be on top of itself. The notion of {\em state-constraint} is very general and has been used in different areas of AI and for different purposes.  If we restrict ourselves to planning, {\em state-constraints} are abstractions for the compact specification of a set of states. For example, {\em state-constraints} in planning allow to specify the set of states where a given action is applicable, the set of states where a given {\em axiom} or {\em derived predicate} holds or the set of states that are considered goal states.

@@

\subsection{Pruning inconsistent action models with {\em schematic mutex}}
For every {\em state-constraint} $\phi\in\Phi$ we could extend $\mathsf{apply_{\xi,\omega}}$ actions with a conditional effect $\{\neg\phi\}\rhd\{mode_{inval}\}$ that checks the consistency of the {\em state-constraints} $\phi$ at every state traversed by a solution to the compiled problem. Cheking arbitrary $\phi$ formulae can however be too expensive for current classical planners. Instead, our approach is to check {\em state-constraints} in the form of {\em schematic mutex}. To implement this checking we add new conditional effects to {\em insert} and {\em apply} actions of the classical planning compilation for the learning of \strips\ action models~\cite{aineto2018learning}. These new conditional effects capture when the programmed model is inconsistent with a {\em schematic mutex} in $\Phi$.

Figure~\ref{fig:ceffects} summarizes the new conditional effects added to the classical planning compilation for the learning of \strips\ action models from {\em schematic mutex}. Next we describe each of them in detail:
\begin{enumerate}
\item[1-3] For every {\em schematic mutex} $\tup{p,q}$ s.t. both $p$ and $q$ belong to $\in{\mathcal I}_{\Psi,\xi}$ a conditional effect is added to the $\mathsf{insertPre_{p,\xi}}$ actions to ban the insertion of two preconditions that are {\em schematic mutex}. Likewise, two conditional effects are added to the $\mathsf{insertEff_{p,\xi}}$ actions to ban the insertion of two positive/negative effects that are {\em schematic mutex}.
\item[4-5] For every {\em schematic mutex} $\tup{p,q}$ s.t. both $p$ and $q$ belong to $\in{\mathcal I}_{\Psi,\xi}$ two conditional effects are added to the $\mathsf{apply_{\xi,\omega}}$ actions to ban positive effects that are inconsistent with an input observation.
\end{enumerate}

\begin{figure}
\begin{footnotesize}
\begin{tabular}{lll}
{\bf ID} & {\bf Action} & {\bf New conditional effect}\\\hline
1&$\mathsf{insertPre_{p,\xi}}$&$\{pre\_q\_\xi\}\rhd\{mode_{inval}\}$\\
2&$\mathsf{insertEff_{p,\xi}}$&$\{pre\_q\_\xi,eff\_q\_\xi,pre\_p\_\xi\}\rhd\{mode_{inval}\}$\\
3&$\mathsf{insertEff_{p,\xi}}$&$\{\neg pre\_q\_\xi,eff\_q\_\xi,\neg pre\_p\_\xi\}\rhd\{mode_{inval}\}$\\
4&$\mathsf{apply_{\xi,\omega}}$&$\{\neg pre\_p\_\xi \wedge eff\_p\_\xi \wedge $\\
&&$q(\omega)\wedge \neg pre\_q\_\xi\}\rhd\{mode_{inval}\}$\\
5&$\mathsf{apply_{\xi,\omega}}$&$\{\neg pre\_p\_\xi \wedge eff\_p\_\xi \wedge $\\
&&$q(\omega)\wedge pre\_q\_\xi \wedge \neg eff\_q\_\xi\}\rhd\{mode_{inval}\}$
\end{tabular}
\end{footnotesize}
	\caption{\small Summary of the new conditional effects added to the classical planning compilation for the learning of \strips\ action models.}
	\label{fig:ceffects}
\end{figure}

The goals of the classical planning problem output by the original compilation are extended with the $\neg mode_{inval}$ literal to validate that only states {\em consistent} with the state constraints defined in $\Phi$ are traversed by the solution plans. The $\neg mode_{inval}$ literal allows us also to define a more compact definition $\mathsf{apply_{\xi,\omega}}$ actions than the one previously proposed by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning} that do not require a disjunction to code the possible preconditions of an action schema.
